---
title: "Project 1"
author: "James Poslusny"
format: html
editor: visual
---

## Project 1

**Acknowledgements:**

I know this section usually comes at the end, but I would be remiss not to express my sincerest gratitude to Dr. Post and TA Gabby Taylor for their flexibility and support whilst I was sick and hospitalized. I am sincerely grateful for the opportunity to catch back up. Thank you. \

**Plan of Attack**

Firstly, we'll create the function for steps 1 and 2. Then the function for step 3. Then the 5&6 function. Then the final, step 4 function. Then we'll put it in the wrapper and take care of the plotting functions later.

**Function for steps 1 and 2**

```{r}
library(dplyr)
library(tidyr)
data_init<- function(link, enrollment = "enrollment"){
  dat<- read.csv(url(link))
  dat<- dat %>% select(Area_name,STCOU,ends_with("D")) %>% 
    rename(area_name = Area_name) %>%
    pivot_longer(
      cols = -c(area_name, STCOU),
      names_to = "area_name",
      values_to = enrollment
    )
  return(dat)
}
```

**Function for step 3**

```{r}
data_infograb<- function(dat, enrollment = "enrollment"){
  dat$year<- substr(dat$enrollment,nchar(dat$enrollment)-2,nchar(dat$enrollment))
  dat$meaurement<- substr(dat$enrollment,1,7)
  return(dat)
}
```

**Function for step 5**

```{r}
data_ifcounty_whatstate<- function(dat){
  dat$state<- substr(dat$area_name, nchar(dat$area_name)-2, nchar(dat$area_name))
  return(dat)
}
```

**Function for step 6**

```{r}
data_ifnotcounty_whatdiv<-function(dat){
  dat$state<- substr(dat$area_name, nchar(dat$area_name)-2, nchar(dat$area_name))
  if(dat$state %in% c("CT","ME","MA","NH","RI","VT")){
    dat$division<- "Division 1"
  } else if (dat$state %in% c("NJ","NY","PA")){
    dat$division<- "Division 2"
  } else if (dat$state %in% c("IL","IN","MI","OH","WI")){
    dat$division<- "Division 3"
  } else if (dat$state %in% c("IA","KS","MN","MO","NE","ND","SD")){
    dat$division<- "Division 4"
  } else if (dat$state %in% c("DE","FL","GA","MD","NC","SC","VA","DC","WV")){
    dat$division<- "Division 5"
  } else if (dat$state %in% c("AL","KY","TN","MS")){
    dat$division<- "Division 6"
  } else if (dat$state %in% c("AK","LA","OK","TX")){
    dat$division<- "Division 7"
  } else if (dat$state %in% c("AZ","CO","ID","MT","NV","NM","UT","WY")){
    dat$division<- "Division 8"
  } else if (dat$state %in% c("AL","CA","HI","OR","WA")){
    dat$division<- "Division 9"
  } else {
    dat$division<- "ERROR"
  }
  return(dat)
}
```

**Function for step 4 (takes output for step 3 and calls 5 and 6, returns two tibbles)**

```{r}
data_4toend<- function(dat){
  county_dat<- dat %>% filter(STCOU == "County")
  noncounty_dat<- dat %>% filter(STCOU != "County")
  class(county_dat)<- c("county",class(county_dat))
  class(noncounty_dat)<- c("state",class(noncounty_dat))
  county_return<- data_ifcounty_whatstate(county_dat)
  noncounty_return<-data_ifnotcounty_whatdiv(noncounty_dat)
  listreturn<- list(county_return,noncounty_return)
  return(listreturn)
}
```

**Creating the wrapper function**

```{r}
data_setup<- function(link, enrollment = "enrollment"){
  result<- data_4toend(data_infograb(data_init(link, enrollment),enrollment))
  #I nested it because my brain understands this better than trying to pipe it
}
```

**Calling it for the first two URLs**

```{r}
OneA<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
OneB<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```

**Creating the combiner function**

```{r}
combine<- function(lista, listb){
  countycombo<- rbind(lista[[1]],listb[[1]])
  noncountycombo<- rbind(lista[[2]],lisb[[2]])
  combolist<- list(countycombo, noncountycombo)
  return(combolist)
}
```

**Calling the combiner function**

```{r}
first_two_combined<- combine(OneA,OneB)
```

**Creating the plot.state function**

```{r}
library(ggplot2)
plot.state<- function(dat, enrollment = "enrollment"){
  dat<- dat %>% filter(division != "ERROR") %>% group_by(year, division) %>%
    summarise(value = mean(get(enrollment)))
  ggplot(dat, aes(x = year,y = value, color = division)) + geom_point()
}
```

**Creating the plot.county function**

```{r}
plot.county<- function(dat, enrollment = "enrollment", state = "NC", performance = "top", how_many = 5){
  dat<- get(dat) %>% filter(division != "ERROR") %>% 
    filter(state == get(state))  %>%
    group_by(area_name) %>%
    summarise(value = mean(get(enrollment)))
  if (get(performance) == "top"){
    order(dat, decreasing = T)
  } else if (get(performance == "bottom")){
    order(dat, descending = F)
  } else {
    return("ERROR: PERFORMANCE CAN ONLY BE top OR bottom")
  }
  
  datforplot<- head(dat, get(how_many))
  ggplot(dataforplot, aes(x = area_name, y = value)) + geom_bar()
    
}
```

**Put it Together**

```{r}
OneA<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
OneB<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
first_two_combined<- combine(OneA,OneB)
plot.state(first_two_combined[[2]])
plot.county(first_two_combined[[1]],state = "NC", performance = "top", how_many = 20)
plot.county(first_two_combined[[1]],state = "SC", performance = "bottom", how_many = 7)
plot.county(first_two_combined[[1]])
plot.county(first_two_combined[[1]],state = "PA", performance = "top", how_many = 8)

#NEW URLS
A<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/PST01a.csv")
B<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/PST01b.csv")
C<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/PST01c.csv")
D<- data_setup("https://www4.stat.ncsu.edu/~online/datasets/PST01d.csv")

AB<- combine(A,B)
CD<- combine(C,D)
ABCD<- combine(AB,CD)
plot.state(ABCD)
plot.county(ABCD, state = "CA", performance = "top", how_many = 15)
plot.county(ABCD, state = "TX", performance = "top", how_many = 4)
plot.county(ABCD)
plot.county(ABCD, state = "NY", performance = "top", how_many = 10)
```
